---
title: "Quickstart"
output: html_document
---

```{r}
knitr::opts_chunk$set(eval = FALSE)
```



# Introduction

**pynterface** is an R interface to Python that provides a persistent
Python environment. It provides a number of convenience
functions to simplify communicating with the Python environment, including
initializing and exiting the Python environment; getting and setting variables; 
accessing documentation; importing modules; and generating R interfaces to 
Python functions. **pynterface** also provides helper interfaces for 
**pandas** and **numpy**, and a generic interface for extending the R interface
to custom Python classes.


# Starting the Python environment

Start a Python environment with `pyConnect`. Because **pynterface** uses the
**pysockr** as a backend, `pyConnect` needs to know what port to use and the 
hostname as well as the path to the Python executable. If you're running 
**pysockr** on a standard client (i.e. not a server or on separate clusters)
the defaults should work just fine. **pynterface** is designed to work with
Python version 2.7 (or higher) or Python version 3.6 (or higher). Use 
`pyInfo` to get information on the connected Python environment and `pyExit`
to shut down the Python environment.

```{r}
# assuming Python is installed in C:/Python27
pyConnect("C:/Python27/python.exe")
# get information on the Python environment
pyInfo()
# shut down the Python environment
pyExit()
```


# Accessing help

You can easily access Python help documents from R using `pyHelp`.

```{r}
pyHelp("str")     # get help on the Python "str" function
pyHelp("pydoc")   # get hep on the pydoc module
```


# Executing Python code

**pynterface** provides a number of functions for executing Python
code from R:

1. `pyexec`: execute Python code without a return value. Useful for
   e.g. import statements.
2. `pyexecp`: execute Python code and print the result.
3. `pyexecg`: execute multiple lines of Python code and retrieve 
   variables.
4. pyExecfile: Compile and execute Python code in a file.

```{r}
pyExec("import os")
pyExecp("os.getcwd()")
pyExecg(c("a = 5", "b = 2", "c = 3"), returnValues = c("a", "c"))

f = tempfile(fileext = ".py")
cat("d = a + b + c", sep = "\n", file = f)
pyExecfile(f)
pyExecp("d")
```


# Getting and Setting Variables

You can also get, set and print variables using the helper functions 
`pyGet`, `pySet` and `pyPrint`:

```{r}
pySet(a = 5)
pyGet('a')
pyPrint('a')
```

`pyGet` and `pySet` use JSON to serialize and parse Python objects via
the **rjson** package on the R side and the **json** module on the 
Python side. This makes it easy to send complex objects back and forth 
between R and Python:

```{r}
pySet(aa = 1:10)
pyGet('aa')
pyPrint('aa')

pySet(dd = data.frame(a = 5, b = "FOO"))

pyExecp("type(dd)")
pyPrint('dd')

pyGet('dd')
```

Objects get converted based on the defaults available to **rjson** 
package and **json** module. Not every Python object can be 
automatically serialized to JSON. Some objects have a method that 
can be called to convert to JSON. For example, **numpy** vectors
can't be automatically converted to JSON strings, but they can be
converted to lists which can then be converted to JSON strings.
**pandas** data frames have a `to_JSON` method for converting to
JSON strings. **pynterface** can use the extensions provided by
**pysockr** for converting **numpy** and **pandas** objects to JSON 
strings for easy reading into R. These extensions can be loaded with 
`useNumpy` and `usePandas`:

```{r}
pyExec("import numpy")
useNumpy()    # add the numpy extension for JSON serializing
usePandas()   # add the pandas extension for JSON serializing
pySet(a = sample(1:10, 5))
pyExec("b = numpy.sort(a)")
pyGet("b")
```

Definitions for custom classes can be loaded with `useClass`.
Defining extensions for custom classes is straightforward; see the 
**pysockr** documentation for more info.


# Generating Interfaces to Python Functions

**pynterface** can automatically generate interfaces to Python 
functions with `pyFunction`:

```{r}
npsort = pyFunction("numpy.sort")
npsort(sample(1:10, 5))
```

Python function interfaces use the undescore character `_` as a
placeholder for the output variable. The `finalizer` argument can
be used to do some final processing of the Python output prior to
JSON serialization and transfer to R.

The `pyImport` function loads a module and attempts to generate an
interface for every function in the module into an R environment:

```{r}
e = new.env()
pyImport("numpy", env = e)
ls(e)
```

# Need more control?

Check out **pysockr** for a low-level, extensible R6 class
interface to Python.
